---
title: "<span class='pres-date'>Week `r params$this_week`: </span><br/><span class='pres-subtitle'>Spatial Data Analysis with R</span> <br/><span class='pres-date'>`r load('../common/course_info.RData'); lstCI[[paste0('class0', params$this_week)]]`</span>"
author: "<span class='pres-title'>R-ArcGIS Bridge</span>"
output: 
  slidy_presentation:
    css: "../common/slidy_styles.css"
    df_print: default
    duration: 0
    font_adjustment: 2
    footer:  !expr load('../common/course_info.RData'); paste("<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span><span class='footer-right'><a href='../index.html'>Course Homepage</a></span><span class='footer-right'>Etherpad:&nbsp<a href='", lstCI$etherpadshort, "' target='_blank'>", lstCI$etherpadshort, "</a></span><span class='footer-subtitle'>Spatial Data Analysis in R</span>")
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
params:
    this_week: 4
---

```{r setup, child = 'slides_setup.Rmd', include=FALSE}
```

```{r load_libraries, include=FALSE}
library(arcgisbinding)
```

# R-ArcGIS Bridge 

## Compatibility & Installation

Requires a licensed ArcGIS desktop version

The R-ArcGIS Bridge works for 

<div class="compact">
- ArcPro 2.0 (native support)
    - Install from ArcPro 
    - [Instruction Video](https://community.esri.com/videos/4136-installing-the-r-arcgis-bridge-for-arcgis-pro-20)
- ArcMap 10.3.1 / ArcPro 1.1
    - [Instructions on GitHub](https://github.com/R-ArcGIS/r-bridge-install)
</div>

## What You Can Do with the Bridge

<div class="compact">
- View metadata before bringing spatial info into R.
- Import spatial layers from disk, or file geodatabases.
- Create R scripts that you run from ArcGIS.
</div>

# Example

```{r loadarcbinding02, cache=TRUE}
library(arcgisbinding)
library(sp)
arc.check_product()

## Create the connection
crimepts_con <- arc.open(path="../exercises/data/SF_Crime.gdb/San_Francisco_Crimes")

## Display info (metadata)
crimepts_con
```

Import to R as "ArcGIS data frame".

```{r loadarcbinding03, cache=TRUE}
crimepts_adf <- arc.select(crimepts_con)
names(crimepts_adf)
nrow(crimepts_adf)
attr(crimepts_adf, "shape")
```

Convert to a <tt>sp</tt> object.

```{r loadarcbinding04, cache=TRUE}
crimepts_spdf <- arc.data2sp(crimepts_adf)
class(crimepts_spdf)
```

Plot.

```{r loadarcbinding05, cache=TRUE}
plot(crimepts_spdf, axes=TRUE, pch=16, cex=0.3)
```

# See also

<div class="compact">
- [arcgisbinding Vignette](https://r-arcgis.github.io/assets/arcgisbinding-vignette.html)
- https://spatialstats.github.io/r-arcgis-bridge/
- [Code samples on GitHub](https://r-arcgis.github.io/)
- ESRI Webinars
</div>

## R-ArcGIS Bridge Exercise

[Analyze Crime Statistics](http://learn.arcgis.com/en/projects/analyze-crime-using-statistics-and-the-r-arcgis-bridge/){target="_blank"}

[Identify an Ecological Niche for African Buffalo](http://learn.arcgis.com/en/projects/identify-an-ecological-niche-for-african-buffalo/){target="_blank"}



