---
title: "<span class='pres-date'>Week `r params$this_week`: </span><br/><span class='pres-subtitle'>Spatial Data Analysis with R</span> <br/><span class='pres-date'>`r load('../common/course_info.RData'); lstCI[[paste0('class0', params$this_week)]]`</span>"
author: "<span class='pres-title'>Raster Data Pt 2: Derived Surfaces</span>"
output: 
  slidy_presentation:
    css: "../common/slidy_styles.css"
    df_print: default
    duration: 0
    font_adjustment: 2
    footer:  !expr load('../common/course_info.RData'); paste("<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span><span class='footer-right'><a href='../index.html'>Course Homepage</a></span><span class='footer-right'>Etherpad:&nbsp<a href='", lstCI$etherpadshort, "' target='_blank'>", lstCI$etherpadshort, "</a></span><span class='footer-subtitle'>Spatial Data Analysis in R</span>")
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
params:
    this_week: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(raster)
library(rgdal)
```


# Working with Pixel Values

First we bring in the elevation data we made last time.

```{r plot_start, cache=TRUE}
load("~/yosemite_bnd.RData")  # yosem_bnd_prj
load("~/yosemite_dem.RData")  # yosem_msk_dem_prj
plot(yosem_msk_dem_prj, asp=1)
plot(yosem_bnd_prj, add=TRUE, lwd=2)
```

There are couple of ways to get pixel values

<tt>myrast[*rows*,*cols*]</tt>

<tt>getValues()</tt> - returns all pixel values as a vector (useful for global operations)

Example: identify steep slopes in Yosemite

```{r hist_slope_vals, cache=TRUE}
yosem_slope <- raster("~/yosemite_slope.tif")
all_slope_vals <- getValues(yosem_slope)
## Slope is computed in radians. To show degrees, multiply by 180.
hist(all_slope_vals * 180, xlab="slope (degrees)", col="grey80", main="Yosemite Slope Distribution")
```

Let's make a mask of high slope values, where we define steep slopes as > 40 degrees.

```{r steep_slope_define, cache=TRUE}
## Create a blank copy of yosem_slope
yosem_slope_steep <- raster(yosem_slope)
hasValues(yosem_slope_steep)

threshhold_deg <- 40
yosem_slope_steep[yosem_slope > (threshhold_deg / 180)] <- 1
yosem_slope_steep
summary(yosem_slope_steep)
plot(yosem_slope)
plot(yosem_slope_steep, col="red", add=TRUE, legend=FALSE)
```

# Extract Pixel Values 

Create some sample points

```{r samp_pts, cache=TRUE}
bbox <- yosem_bnd_prj@bbox
num_pts <- 20
x <- runif(min=bbox[1,1], max=bbox[1,2], num_pts)
y <- runif(min=bbox[2,1], max=bbox[2,2], num_pts)
yosem_pts <- SpatialPoints(coords=data.frame(x=x,y=y), proj4string = yosem_bnd_prj@proj4string)
plot(yosem_bnd_prj, axes=T, asp=1)
plot(yosem_pts, cex=2, pch=16, col="red", add=TRUE)
```

To extract values from a raster for where features overlap, use <tt>extract()</tt>.

```{r}
plot(yosem_msk_dem_prj)
plot(yosem_pts, cex=2, pch=16, col="red", add=TRUE)
pt_elev <- extract(yosem_msk_dem_prj, yosem_pts)
pt_elev
```

# Rastering Vector Data

To do analysis, we often have to convert vector features to raster format.

Syntax: <tt>rasterize(*spgeom*, *rast*, *field*)</tt>

```{r rasterize_ex, cache=TRUE}
yosem_pts_rst <- rasterize(yosem_pts, yosem_slope, field=1)
x <- getValues(yosem_pts_rst)
table(x)
```

## Create a distance surface

```{r create_dist, cache=TRUE}
yosem_dist2pts <- distance(yosem_pts_rst)
plot(yosem_dist2pts, main="Distance to Random Points")
plot(yosem_bnd_prj, add=TRUE, lwd=2)
```

# Raster Algebra

Raster algebra is where you add, subtract, multiply, etc. two or more rasters together. 

- The rasters should have the same extent and pixel size

- The operation is evaluated on a pixel-by-pixel basis.

- We can use raster algebra for siting analyses that aim to find the locations that meet multiple criteria. 


Example: Find all locations in Yosemite that are within 1 mile of our sample points and all on slopes < 20 degrees. 

```{r plot_rast01, cache=TRUE}
## Create a raster of the suitable distances
## All pixels will be NA except for those that meet the criteria
# 1 mile = 1609.34 meters
pts_close_rst <- raster(yosem_dist2pts)
dist_vals <- getValues(yosem_dist2pts)
pts_close_rst[ dist_vals < 1609.34  ] <- 1

## Create a raster of suitable flat areas
flat_areas_rst <- raster(yosem_slope)
slope_vals <- getValues(yosem_slope)
flat_areas_rst[ slope_vals < (20/180) ] <- 1

## Multiple the two rasters to find those cells where both conditions are met
best_locations <- flat_areas_rst * pts_close_rst
image(best_locations, col="red")
plot(yosem_bnd_prj, add=TRUE, lwd=2)
```

# Next Up

[Exercise 11: Derived Data](../exercises/ex11_raster02.html){target="_blank"}

[Geocoding](part03e_geocoding.html)


