---
title: "<span class='pres-subtitle'>Spatial Data Analysis with R</span>"
author: "<span class='pres-title'>Week 2: Intro</span>"
date: "`r load('../common/course_info.RData'); lstCI$class02`"
output: 
  slidy_presentation:
    css: "../common/slidy_styles.css"
    df_print: default
    duration: 0
    font_adjustment: 2
    footer: "<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span> 
    <span class='footer-right'><a href='../index.html'>Course Homepage</a></span>
    <span class='footer-right'>Etherpad: <a href='https://is.gd/rspatial_baygeo_s18'>https://is.gd/rspatial_baygeo_s18</a></span>
    <span class='footer-subtitle'>Spatial Data Analysis in R: Week 2</span>"
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
params:
   this_week: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Course Outline

```{r test-main, child = 'course_outline.Rmd'}
```

# Review of Last Time

## R Syntax

- case sensitive
- console - where commands get executed
- script (source) window - can save commands to a files

## Operators

### Math

<tt>+    -    *    %    :    %%     %/%</tt>

### Comparison

<tt>==     !=</tt>

<tt>>  >=   <   <=   </tt>

<tt>%in%</tt>   

## Functions

### Math

<tt>seq(), c(), sample(), runif(), log(), mean(), sum(), table()</tt>

### Finding out about objects

<tt>class(), summary()</tt>

### Controlling the R environment

<tt>setwd(), getwd()</tt>

## Variables

<tt>x = "san francisco"</tt>  

<tt>x <- "san francisco"</tt>

## Data Classes

vector  
   character  
   numeric  
   Boolean  

## Vector Functions / Element-wise Vector Functions

<tt>x <- 1:3</tt>
<tt>x ^ 2</tt>
<tt>log(x)</tt>

## Filtering with Square Brackets

```{r}
x <- 4:6  
idx <- c(1,2)
x[idx]
 

idx <- c(T,T,F)
x[idx]

idx <- x > 4
x[idx]
```

