---
title: "<span class='pres-date'>Week `r params$this_week`: </span><br/><span class='pres-subtitle'>Spatial Data Analysis with R</span> <br/><span class='pres-date'>`r load('../common/course_info.RData'); lstCI[[paste0('class0', params$this_week)]]`</span>"
author: "<span class='pres-title'>R Syntax</span>"
output: 
  slidy_presentation:
    css: "../common/slidy_styles.css"
    df_print: default
    duration: 0
    font_adjustment: 2
    footer:  !expr load('../common/course_info.RData'); paste("<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span><span class='footer-right'><a href='../index.html'>Course Homepage</a></span><span class='footer-right'>Etherpad:&nbsp<a href='", lstCI$etherpadshort, "' target='_blank'>", lstCI$etherpadshort, "</a></span><span class='footer-subtitle'>Spatial Data Analysis in R</span>")
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
params:
    this_week: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R as a calculator

```{r computing,echo=TRUE,results='markup'}
2 + 2 # add numbers
2 * pi # multiply by a constant
7 + runif(1) # add a random number
3^4 # powers
sqrt(4^4) # functions
log(10)
log(100, base = 10)
23 %/% 2 
23 %% 2

# scientific notation
5000000000 * 1000
5e9 * 1e3
```

Think of a mathematical operation you need - can you guess how to do it in R?

# Assigning values to R objects

A key action in R is to store values in the form of variables (also called objects). We can use the '<tt>=</tt>' or '<tt><-</tt>' operators to do this.

```{r}
y <- 3^3
```

Note when you assign a value to a variable, R doesn't print the result to the console. To see the value of a variable, simply type it by itself.

```{r}
y
```

Pretty much anything can be saved to a variable. Here we save some text and dates.

```{r}
flv <- "vanilla"
flv

start_date <- Sys.Date()
start_date
```

## Naming Variables

The rules for naming variables are pretty flexible. You can use numbers, letters, and most special characters. There are a couple of rules to take note of:

<div class="indented2">
- variable names can't start with a number (e.g., *3pieces*)  
- hyphens and spaces are not allowed
</div>

__Naming Styles__

There are a handful of popular  [naming styles](https://www.r-bloggers.com/consistent-naming-conventions-in-r/){target="_blank"}. Pick one that you like, and be consistent! 

```{r, echo=FALSE}
library(kableExtra)
naming_styles_vec <- c("alllowercase", "adjustcolor", "period.separated", "shoe.size", "underscore_separated", "numeric_version", "lowerCamelCase", "addTaskCallback", "UpperCamelCase", "SignatureMethod")
naming_styles_mat <- matrix(naming_styles_vec, byrow=TRUE, ncol=2, dimnames=list(1:(length(naming_styles_vec)/2), c("Style", "Example")))
knitr::kable(naming_styles_mat, format = "html", table.attr='class="tbl_compact"') %>% column_spec(column=2, monospace = TRUE)
```

# R Studio Shortcuts

If you're starting to type something you've typed before, or the long name of an R object or function, STOP!  You likely don't need to type all of that.

- Tab completion
- Command history 
    * up/down arrows
    * Ctrl-{up arrow} or Command-{up arrow}
- RStudio: select a line or block for execution
- Put your code in a file and use `source()`. For example: `source('myRcodeFile.R')`

**Question**: Are there other tricks that anyone knows of?

# Vectors in R

The most basic form of an R object is a vector. In fact, individual (scalar) values are vectors of length one. 

We can concatenate values into a vector with `c()`.

```{r}
## numeric vector
nums <- c(1.1, 3, -5.7)
devs <- rnorm(5)
devs

## integer vector
ints <- c(1L, 5L, -3L) # force storage as integer not decimal number
## 'L' is for 'long integer' (historical)

idevs <- sample(ints, 100, replace = TRUE)

## character vector
chars <- c('hi', 'hallo', "mother's", 'father\'s', 
   "She said, 'hi'", "He said, \"hi\"" )
chars
cat(chars, sep = "\n")

## logical vector
bools <- c(TRUE, FALSE, TRUE)
bools
```

# Sequences

Sequences of numbers are useful for all sorts of things such as setting up loops and data wrangling. The <tt>seq()</tt> function has a full set of arguments to make sequences, or you can use the <tt>:</tt> operator if you just need a set of sequential integers.    

```{r}
mySeq <- 1:6
mySeq

myOtherSeq <- seq(1, 11, by = 2)
myOtherSeq
length(myOtherSeq)

fours <- rep(4, 6)
fours

## This is a comment: here is an example of non-numeric data
depts <- c('espm', 'pmb', 'stats')
depts
```

# Random Numbers

R has a number of random number generators. We've already seen the runif() function which generates continuous random numbers along a uniform distribution. The rnorm() function is similar but it generates normally distributed numbers.

```{r}
x <- round(10 * runif(1000))
table(x)

x <- rnorm(1000)
hist(x)
```


# Working with indices and subsets

Vectors can be incredibly convenient and greatly simplify your code. However you often don't want all of the elements in a vector. Fortunately R provides a convenient way to extract specific elements from a vector using square-bracket notation. There are three ways you use square brackets.

## Option 1: pass a vector of integers

If you pass a vector of integers in the square brackets, R will take them to be the indices of the element(s) you want returned.

```{r}
vals <- seq(2, 12, by = 2)
vals

vals[3]
vals[3:5]
vals[c(1, 3, 6)]
```

Challenge question: what happens if you pass a negative integer? [<span id="qtn_tismynn" class="showhide" style="cursor: pointer; color:blue;" onclick="showHide('tismynn');return false;">Ans</span>].

<div id="tismynn" class="answer-code" style="margin-left:1em;">
Negative integers tell R to *omit* those elements.

```{r}
x <- LETTERS[1:5]
x[-2]
```
</div>

## Option 2: Pass a vector of logicals (Booleans)

The second way to extract element(s) from a vector is to pass a vector of TRUE/FALSE values in the square brackets. R will return the elements corresponding to the TRUE values (assuming the TRUE/FALSE values are in the same order as the elements of the vector).

```{r}
vals
vals[c(T,F,F,F,T,T)]
```

## Option 3: Pass an expression that returns a vector of Booleans

The is really the same as option #2, but instead of instead of creating the vector of Booleans as a separate object you enter an expression generates them on-the-fly.

```{r}
x <- 1:10
x[ x > 5 ]
x[ x %% 2 == 0 ]
```

# Gotcha's

## R  is case sensitive

```{r, echo=TRUE, message=TRUE, warning=TRUE}
"Hello" == "Hello"
"GoodBye" == "goodbye"
mean(1:5)
## Mean(1:5)  - will throw an error: could not find function "Mean"
```

# Next-Up

[Exercise 1. R Syntax](../exercises/ex01_basicr.html){target="_blank"}

[Working With Packages](part01c_packages.html)
