---
title: "<span class='pres-subtitle'>Spatial Data Analysis with R</span>"
author: "<span class='pres-title'>Week 1: Packages and R Syntax</span>"
output: 
  slidy_presentation:
    css: slidy_styles.css
    df_print: default
    duration: 0
    font_adjustment: 2
    footer: "<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span> 
    <span class='footer-right'><a href='../index.html'>Course Homepage</a></span>
    <span class='footer-right'>Etherpad: <a href='https://is.gd/rspatial_baygeo_s18'>https://is.gd/rspatial_baygeo_s18</a></span>
    <span class='footer-subtitle'>Spatial Data Analysis in R: Week 1</span>"
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: header.html
params:
   this_week: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Packages

<div class="incremental"><div><hr class="hrdivider">
__Finding Packages__

CRAN Task Views: https://cran.r-project.org/web/views/Spatial.html

Forums, Blogs, Google

</div></div>


<div class="incremental"><div><hr class="hrdivider">
__Installing and Downloading__

Install (downloading)

_From CRAN repositories_

```{r, eval=FALSE, echo=TRUE}
install.packages("raster", dependencies = TRUE, repos="http://cran.cnr.berkeley.edu/")
```

_From GitHub_

```{r, eval=FALSE, echo=TRUE}
library(devtools)
install_github("sidtraits", "Koalha")
```

_Installation Errors_

```{r, echo=TRUE}
install.packages("Roster", repos="http://cran.cnr.berkeley.edu/")
```

Most of the time, a warning like this means the package doesn't exist at all.

_Load into Memory_

```{r, echo=TRUE}
library("raster")
```

To unload:

```{r, echo=TRUE, eval=FALSE}
detach("package:raster")
```

To update:

```{r, echo=TRUE, eval=FALSE}
## update all installed packages
update.packages() 

## update a single package
update.packages(oldPkgs="raster")
```

</div></div>

<div class="incremental"><div><hr class="hrdivider">
__What's in a package?__

<table>
<tr><td>Index of all help pages</td><td><tt>help(package="maptools")</tt></td></tr>

<tr><td>List of Vignettes</td><td><tt>browseVignettes(package = "raster")</tt></td></tr>

<tr><td>Bundled datasets</td><td><tt>data(package="myPackge")</tt></td></tr>

<tr><td>Installation directory</td><td><tt>system.file(package = "maptools")</tt></td></tr>

<tr><td>Bundled demos</td><td><tt>demo(package="sp")</tt></td></tr>

<tr><td>Run a demo</td><td><tt>demo(webmap)</tt></td></tr>
</table>
</div></div>

# Exercise 01

<a href="../exercises/part01_exercise1_packages.html" target="_blank">Exercise 1: Packages</a>

# R Syntax 

# R as a calculator

```{r computing,echo=TRUE,results='markup'}
2 + 2 # add numbers
2 * pi # multiply by a constant
7 + runif(1) # add a random number
3^4 # powers
sqrt(4^4) # functions
log(10)
log(100, base = 10)
23 %/% 2 
23 %% 2

# scientific notation
5000000000 * 1000
5e9 * 1e3
```

Think of a mathematical operation you need - can you guess how to do it in R?

# Assigning values to R objects

A key action in R is to store values in the form of R objects, and to examine the value of R objects.

```{r}
val <- 3
val
print(val)

Val <- 7 # case-sensitive!
print(c(val, Val))
```

We can work with (and store) sequences and repetitions
```{r}
mySeq <- 1:6
mySeq

myOtherSeq <- seq(1.1, 11.1, by = 2)
myOtherSeq
length(myOtherSeq)

fours <- rep(4, 6)
fours

## This is a comment: here is an example of non-numeric data
depts <- c('espm', 'pmb', 'stats')
depts
```

If we don't assign the output of a command to an object, we haven't saved it for later use.

R gives us a lot of flexibility (within certain rules) for assigning to (parts of) objects from (parts of) other objects.

# How to be [lazy](http://dilbert.com/strips/comic/2005-05-29/)

If you're starting to type something you've typed before, or the long name of an R object or function, STOP!  You likely don't need to type all of that.

- Tab completion
- Command history 
    * up/down arrows
    * Ctrl-{up arrow} or Command-{up arrow}
- RStudio: select a line or block for execution
- Put your code in a file and use `source()`. For example: `source('myRcodeFile.R')`

**Question**: Are there other tricks that anyone knows of?

# Vectors in R

The most basic form of an R object is a vector. In fact, individual (scalar) values are vectors of length one. 

We can concatenate values into a vector with `c()`.

```{r}
## numeric vector
nums <- c(1.1, 3, -5.7)
devs <- rnorm(5)
devs

## integer vector
ints <- c(1L, 5L, -3L) # force storage as integer not decimal number
## 'L' is for 'long integer' (historical)

idevs <- sample(ints, 100, replace = TRUE)

## character vector
chars <- c('hi', 'hallo', "mother's", 'father\'s', 
   "She said, 'hi'", "He said, \"hi\"" )
chars
cat(chars, sep = "\n")

## logical vector
bools <- c(TRUE, FALSE, TRUE)
bools
```

# Working with indices and subsets

```{r}
vals <- seq(2, 12, by = 2)
vals

vals[3]
vals[3:5]
vals[c(1, 3, 6)]
vals[-c(1, 3, 6)]
vals[c(rep(TRUE, 3), rep(FALSE, 2), TRUE)]
```

<div class="incremental"><div><hr class="hrdivider">
__R is case sensitive__

```{r, echo=TRUE, message=TRUE, warning=TRUE}
"Hello" == "Hello"
"GoodBye" == "goodbye"
mean(1:5)
## Mean(1:5)  - will throw an error: could not find function "Mean"
```
</div></div>

# Exercise

[Exercise 2 (do questions 1-10)](../exercises/part01_exercise2_basicr.html)




