---
title: "<span class='pres-date'>Week `r params$this_week`: </span><br/><span class='pres-subtitle'>Spatial Data Analysis with R</span> <br/><span class='pres-date'>`r load('../common/course_info.RData'); lstCI[[paste0('class0', params$this_week)]]`</span>"
author: "<span class='pres-title'>Control Structures: Loops and Branches</span>"
output: 
  slidy_presentation:
    css: "../common/slidy_styles.css"
    df_print: default
    duration: 0
    font_adjustment: 2
    footer:  !expr load('../common/course_info.RData'); paste("<span class='footer-right'><a href='acknowledgements.html'>Acknowledgements</a></span><span class='footer-right'><a href='../index.html'>Course Homepage</a></span><span class='footer-right'>Etherpad:&nbsp<a href='", lstCI$etherpadshort, "' target='_blank'>", lstCI$etherpadshort, "</a></span><span class='footer-subtitle'>Spatial Data Analysis in R</span>")
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
params:
    this_week: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

---

# Course Outline

```{r test-main, child = 'course_outline.Rmd'}
```

# Branching with <tt>if</tt>, <tt>else</tt>, and <tt>ifelse</tt>

<tt>if</tt> and <tt>else</tt> statements are control structures that let you control how a code should execute based on a set of pre-established conditions.

Syntax:

<tt>if (condition) {statement} else {other statement}</tt>

```{r}
x <- 7
if (x > 7) {
  print(x)
} else {
  print("NOT BIG ENOUGH!!")
}
```

*Notes*

<div class="indented1">
- the conditional expression after the <tt>if</tt> statement much be in (parentheses)  
- Use {curly braces} to define chunks of code   
- indentation is not mandatory, but makes the code more legible   
</div>

__<tt>ifelse()</tt>__

A variabtion of the <tt>if-then</tt> statements is the __<tt>ifelse()</tt>__ function. This function is handy for recoding data.

Syntax:

<tt>ifelse(test, yes, no)</tt>


```{r, eval=FALSE}
animal <- ifelse(animal=="cat", 1, 0)       
```


# Loops

- Use a __for__ loop if you know in advance how many passes  
- use a __while__ loop if you have to do a check to see if you need to continue going through the loop  

## For loops

You generally loop a variable like *i* over a vector of values to control the loop.

```{r}
for (i in 1:10){
  print(i^2)
}
```

An example where the loop variable is used as an index.

```{r}
for (i in 1:10) {
  print(state.name[i])
}
```

Loops can be combined with if-else statements

```{r}
x <- 1:10
for (i in 1:10) {
    if(x[i] >= 7) { 
        print(x[i])
    } else {                     
      print("NOT BIG ENOUGH!!")
    }
}
```

## Keeping Track of a Variable

A common practice is to 


Loop through a data frame




## While Loops




# Next Up

[Exercise 8: Loops and Branches](exercises/ex08_loops-branches.html){target="_blank"}

[Spatial Queries](part03b_spatial-queries.html)


