---
title: "<span class='ex-title'>Spatial Analysis with R:</span><br/><span class='ex-subtitle'>Exercise 1: Basic R Operations</span>"
date: "`r load('../common/course_info.RData'); lstCI$course`"
output: 
  html_document: 
    css: "../common/slidy_styles.css"
    self_contained: no
    lib_dir: lib
    smart: no
    includes:
      in_header: "../common/header.html"
      after_body: bottom_links.html
---

```{r setup, child = '../common/exercises_setup.Rmd', include=FALSE}
```

```{r preload_libraries, include=FALSE}

```

### Instructions

- If you haven't already, install [R](https://cran.r-project.org/) and [RStudio](https://www.rstudio.com/products/rstudio/download/). 

- Download the <a href="`r lstCI$githubrepo`" target="_blank">GitHub repo</a>. 

- Create a new R script in RStudio (File &rarr; New File &rarr; R Script). 

- Copy the questions below (all of which begin with the _#_ character so RStudio will interpret them as a comment) into your script. Under each question, write R code. Run each line of R code as you enter it (press the Run button or ctrl+enter) to make sure it works.

- Remember to save your script periodically to avoid losing work if RStudio crashes.

- If you get stuck, click on the 'hint' link, or the question number to see the answer. Click again to hide.

- If you have any questions or problems, please feel free to email the instructor.

<hr class="hrdivider">
<div class="copy-button"><button onclick="CopyToClipboard('qtn_div');return false;">Copy questions</button></div>
<div id="qtn_div" class="ex-qtn">

`r hashtitle("BASIC ARITHMETIC, VECTORS, VARIABLES, AND RANDOM NUMBERS")`

<a href="#" class="showhidelink" onclick="showHide('q1');return false;">\# `r auton()`</a> Write an expression that will return the area of a circle with radius=2. <a href="#" class="showhidehint" onclick="showHide('pmnmny');return false;"></a>  

<div id="pmnmny" class="hint">*area = pi times radius squared*</div>


<div id="q1" class="answer-code">
```{r}
# circle area = pi x r squared
pi * 2 ^ 2
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q2');return false;">\# `r auton()`</a> `r chop("Create a variable named _x_, and assign it a numeric value. Then write an expression that returns the diameter of a circle with radius = _x_.")` <a href="#" class="showhidehint" onclick="showHide('satbcl');return false;"></a>  

<div id="satbcl" class="hint">*diameter = 2 times radius*</div>


<div id="q2" class="answer-code">
```{r}
x = 3
# circle diameter = 2 x pi x r
2 * pi * x
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q3');return false;">\# `r auton()`</a> Write an expression that returns a sequence of integers from 1 to 10.

<div id="q3" class="answer-code">
```{r}
1:10
# or:
seq(1, 10)
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q4');return false;">\# `r auton()`</a> Write an expression that returns all the _even_ integers from 1 to 20.

<div id="q4" class="answer-code">
```{r}
1:10 * 2
# or:
seq(from=2, to=20, by=2)
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q5');return false;">\# `r auton()`</a> Write an expression that returns ten uniformly distributed random numbers between 0 and 100. <a href="#" class="showhidehint" onclick="showHide('rowygl');return false;"></a>  

<div id="rowygl" class="hint">see <tt>runif()</tt></div>


<div id="q5" class="answer-code">
```{r}
## runif() returns random numbers from 0 to 1, so if we multiply those values by 100 the 
## result will be numbers from 0 to 100.
runif(n=10) * 100

## If you prefer integers, the sample() function would work also
sample(1:100, size=10)
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q6');return false;">\# `r auton()`</a> `r chop("Create a variable called _y_, and assign it ten randomly selected valuesof TRUE or FALSE (equal weights). Find out how many elements are TRUE.")`  

<div id="q6" class="answer-code">
```{r}
y = sample(c(TRUE,FALSE), size=10, replace=TRUE)

## There are a few ways to count up the number of 'TRUE's
summary(y)
sum(y)
table(y)
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q7');return false;">\# `r auton()`</a> Create a character variable called _h_, and assign it the individual lowercase letters of the alphabet.

<div id="q7" class="answer-code">
```{r}
## Easist is to use a built-in constant:
h <- letters
h

## Simple also works:
h <- c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y", "z")

## *NOT* correct (do you see why?)
## h <- "abcdefghijklmnopqrstuvwxyz"

## Fancy:
h <- strsplit(rawToChar(as.raw(97:122)), split=NULL)[[1]]

```
</div>

`r hashtitle("INDEXING AND FILTERING")`

<a href="#" class="showhidelink" onclick="showHide('q8');return false;">\# `r auton()`</a> Using the variable _h_ you made above, write an expression that returns the first five letters of the alphabet.

<div id="q8" class="answer-code">
```{r}
h[1:5]
# or:
head(h, n=5)

```
</div>

<a href="#" class="showhidelink" onclick="showHide('q9');return false;">\# `r auton()`</a> Write an expression that returns only the vowels.

<div id="q9" class="answer-code">
```{r}
h[c(1,5,9,15,21)]
```
</div>

<a href="#" class="showhidelink" onclick="showHide('q10');return false;">\# `r auton()`</a> `r chop("Create a variable called _x_ and assign it a sequence of integers from 1 to 10. Then change the 5th element to -99. Next, add 100 to the last five elements. Finally, find the average (mean) of your vector.")`  

<div id="q10" class="answer-code">
```{r}
x <- 1:10
x[5] <- -99
x[6:10] <- x[6:10] + 100
x
mean(x)
```
</div>

</div>