---
title: 'Exercise 3: Working with Two-Dimensional Data'
output:
  html_notebook:
    code_folding: hide
    theme: yeti
    toc: yes
    toc_depth: 4
---

## USEFUL FUNCTIONS

Some useful functions you may want to use in this exercise include:

class()  
data()  
head()  
hist()  
mean()  
names()  
nrow()  
read.csv()  
sd()  
sort()    
summary()  
table()  
tolower()  
unique()  

## SORTING VECTORS

__[1] Generate 500 random numbers uniformly distributed between 0 and 100. Find the largest five and
smallest five values. Create a historgram. Bonus: repeat with a normal distribution.__ 

```{r}

```

## BUNDLED DATA

Many R packages come with bundled data sets. The data() function will list available datasets from the datasets package. You generally don’t have to ‘load’ a bundled dataset into memory, R will automatically bring them in when needed. Most datasets also have a help page that explain their contents, that you can view in the same way you view help page for a function.

__[2] Explore the 'iris' dataset: What type of object (class) is it? What columns does it contain? What do the numbers represent?__

```{r}

```

## IMPORTING CSV FILES

The <tt>raster</tt> package comes with a csv file called *countries.csv*. To find out where it is located on your hard drive, run:  

`system.file("external/countries.csv", package="raster")`


__[4] Import _countries.csv_ as a dataframe and save it as countriesDF. Use the <tt>summary()</tt> function to see what each column contains.__

```{r}

```

## TWO-DIMENSIONAL DATA STRUCTURES

The main two-dimensional data structures in R are matrices and data frames. The big differences between the two are that:

i) all the elements of matrices have to be the same data type (usually numeric) whereas each column of a dataframe can be a different data type

ii) the columns of a data frame are usually named, and you can reference them by name

__[5] Display the column names of the countriesDF you just imported. Then make them all lowercase (hint: tolower() returns a lowercase version of its argument).__

```{r}

```

__[6] How many rows are there in the iris data frame?__

```{r}

```

__[7] Write an expression that returns the names of the species in iris.__

```{r}

```

__[8] Write an expression that returns all columns from iris but only the rows for the species virginica.__

```{r}

```

__[9] What is the mean, standard deviation, and quartiles of the petal length for the species _setosa_?__

```{r}

```

__[10] Using the countries data frame you imported earlier, how many countries are there on the continent of Africa?__

```{r}

```

## BASIC PLOTTING

__[11] Create a histogram of state.area (area of the 50 US states in square miles). Investigate what the `breaks` argument does.__ 

```{r}

```

__[12] Create a scatterplot of the sepal length vs petal length from iris. Does it look like they’re correlated? Bonus: create the same plot using ggplot2.__

```{r}

```


