---
title: "<span class='ex-title'>Spatial Analysis with R:</span><br/><span class='ex-subtitle'>Exercise 1: Finding and Installing Packages</span>"
date: "Spring, 2018"
output: 
  html_document: 
    css: slidy_styles.css
    includes:
      in_header: header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing Packages

The key to working effectively in R is finding and installing the right packages. For spatial data, some of the most common and useful packages include:

* __sp__: classes and methods for spatial data
* __rgdal__: importing and transforming different spatial data types and providing access to proj4 library for working with map projections and coordinate systems
* __rgeos__: for spatial operations and queries on geometric objects
* __ggmap__: extends ggplot2 to provide access to online basemaps (e.g. Google Maps, OpenStreetMaps) and online services like the Google Geocoder
* __leaflet__: for creating interactive web maps
* __RColorBrewer__: for selecting predefined color palettes.
* __raster__: functions for working with raster data

1. To ensure you have these packages, run the following lines of code in R. Hint: start a new R script, can copy-paste the lines below, then select all lines and choose 'Run'.

```{r install_packages, echo=TRUE, results='hide', message=FALSE}
## Make a character vector with the require package names
required.pkg <- c("sp", "rgdal", "rgeos", "leaflet", "ggmap", "maptools", "RColorBrewer", "raster", "maps", "zipcode")

## Find out which ones are not already installed
pkgs.not.installed <- required.pkg[!sapply(required.pkg, function(p) require(p, character.only=T))]

## Install any packages needed
if (length(pkgs.not.installed) > 0) install.packages(pkgs.not.installed, dependencies=TRUE, repos="http://cran.cnr.berkeley.edu/")

# (Re)load all libraries at once
lapply(required.pkg, library, character.only = TRUE)
```

2. Run the 'gallery' demo from the <tt>sp</tt> package. <a href="#" class="showhidelink" onclick="showHide('q2');return false;">answer</a>

<div id="q2" class="answer-code">
``` {r, eval=FALSE}
library(gallery)  
demo(gallery)
```
</div>

3. What datasets are bundled with the <tt>maps</tt> package? 
<a href="#" class="showhidelink" onclick="showHide('q3');return false;">answer</a>

<div id="q3" class="answer-code">
``` {r, eval=FALSE}
data(package="maps")
```
</div>


4. What type of object is the 'world.cities' from the <tt>maps</tt> package?
<a href="#" class="showhidelink" onclick="showHide('q4');return false;">answer</a>

<div id="q4" class="answer-code">
``` {r, eval=FALSE}
class(world.cities)  
```
</div>

5. Open the Vignette for the <tt>raster</tt> package
<a href="#" class="showhidelink" onclick="showHide('q5');return false;">answer</a>

<div id="q5" class="answer-code">
``` {r, eval=FALSE}
vignette("Raster", package="raster")
```
</div>

